\jaxcosmo: An End-to-End Differentiable and GPU Accelerated Cosmology Library

### 2022-08-01 paper_OJA.tex
fix Dark Energy Task force reference

\date{\today}

@@

Fisher forecasts are also commonly used in survey and analysis strategy strategy, where running a full posterior analysis for each possible choice would be unfeasible. The inverse area of a projection of a Fisher matrix in parameters of interest can be used as a metric for survey constraining power, such as in the Dark Energy Task Force report \citep{detf}.

>>>>

%\date{\today}

@@

Fisher forecasts are also commonly used in survey and analysis strategy strategy, where running a full posterior analysis for each possible choice would be unfeasible. The inverse area of a projection of a Fisher matrix in parameters of interest can be used as a metric for survey constraining power, such as in the Dark Energy Task Force report \citep{2006astro.ph..9591A}.

### 2022-08-01 paper_deprecated.tex
fix Dark Energy Task force reference

\PassOptionsToPackage{usenames,dvipsnames}{xcolour}
\documentclass[twocolumn,twocolappendix,nofootinbib,iop]{openjournal}

% Add your own macros here:
\pdfoutput=1 %for arXiv submission
\usepackage{amsmath,amssymb,amstext}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage{ae,aecompl}
\usepackage[utf8]{inputenc}
%\usepackage{hyperref}
\usepackage[colorlinks,allcolors=blue]{hyperref}
\usepackage[figure,figure*]{hypcap}
\usepackage{natbib}
\usepackage{url}
\usepackage{mdwlist}
\usepackage{multirow}
\urlstyle{same}
\usepackage{lineno}
\usepackage{fontawesome}

%%%% usepackage{minted}
\usepackage{xcolor}
\definecolor{maroon}{cmyk}{0, 0.87, 0.68, 0.32}
\definecolor{halfgray}{gray}{0.55}
\definecolor{ipython_frame}{RGB}{207, 207, 207}
\definecolor{ipython_bg}{RGB}{247, 247, 247}
\definecolor{ipython_red}{RGB}{186, 33, 33}
\definecolor{ipython_green}{RGB}{0, 128, 0}
\definecolor{ipython_cyan}{RGB}{64, 128, 128}
\definecolor{ipython_purple}{RGB}{170, 34, 255}

\usepackage{listings}

\lstset{
    breaklines=true,
    %
    extendedchars=true,
    literate=
    {á}{{\'a}}1 {é}{{\'e}}1 {í}{{\'i}}1 {ó}{{\'o}}1 {ú}{{\'u}}1
    {Á}{{\'A}}1 {É}{{\'E}}1 {Í}{{\'I}}1 {Ó}{{\'O}}1 {Ú}{{\'U}}1
    {à}{{\`a}}1 {è}{{\`e}}1 {ì}{{\`i}}1 {ò}{{\`o}}1 {ù}{{\`u}}1
    {À}{{\`A}}1 {È}{{\'E}}1 {Ì}{{\`I}}1 {Ò}{{\`O}}1 {Ù}{{\`U}}1
    {ä}{{\"a}}1 {ë}{{\"e}}1 {ï}{{\"i}}1 {ö}{{\"o}}1 {ü}{{\"u}}1
    {Ä}{{\"A}}1 {Ë}{{\"E}}1 {Ï}{{\"I}}1 {Ö}{{\"O}}1 {Ü}{{\"U}}1
    {â}{{\^a}}1 {ê}{{\^e}}1 {î}{{\^i}}1 {ô}{{\^o}}1 {û}{{\^u}}1
    {Â}{{\^A}}1 {Ê}{{\^E}}1 {Î}{{\^I}}1 {Ô}{{\^O}}1 {Û}{{\^U}}1
    {œ}{{\oe}}1 {Œ}{{\OE}}1 {æ}{{\ae}}1 {Æ}{{\AE}}1 {ß}{{\ss}}1
    {ç}{{\c c}}1 {Ç}{{\c C}}1 {ø}{{\o}}1 {å}{{\r a}}1 {Å}{{\r A}}1
    {€}{{\EUR}}1 {£}{{\pounds}}1
}

%%
%% Python definition (c) 1998 Michael Weber
%% Additional definitions (2013) Alexis Dimitriadis
%% modified by me (should not have empty lines)
%%
\lstdefinelanguage{iPython}{
    morekeywords={access,and,break,class,continue,def,del,elif,else,except,exec,finally,for,from,global,if,import,in,is,lambda,not,or,pass,print,raise,return,try,while},%
    %
    % Built-ins
    morekeywords=[2]{abs,all,any,basestring,bin,bool,bytearray,callable,chr,classmethod,cmp,compile,complex,delattr,dict,dir,divmod,enumerate,eval,execfile,file,filter,float,format,frozenset,getattr,globals,hasattr,hash,help,hex,id,input,int,isinstance,issubclass,iter,len,list,locals,long,map,max,memoryview,min,next,object,oct,open,ord,pow,property,range,raw_input,reduce,reload,repr,reversed,round,set,setattr,slice,sorted,staticmethod,str,sum,super,tuple,type,unichr,unicode,vars,xrange,zip,apply,buffer,coerce,intern},%
    %
    sensitive=true,%
    morecomment=[l]\#,%
    morestring=[b]',%
    morestring=[b]",%
    %
    morestring=[s]{'''}{'''},% used for documentation text (mulitiline strings)
    morestring=[s]{"""}{"""},% added by Philipp Matthias Hahn
    %
    morestring=[s]{r'}{'},% `raw' strings
    morestring=[s]{r"}{"},%
    morestring=[s]{r'''}{'''},%
    morestring=[s]{r"""}{"""},%
    morestring=[s]{u'}{'},% unicode strings
    morestring=[s]{u"}{"},%
    morestring=[s]{u'''}{'''},%
    morestring=[s]{u"""}{"""},%
    %
    % {replace}{replacement}{lenght of replace}
    % *{-}{-}{1} will not replace in comments and so on
    literate=
    {á}{{\'a}}1 {é}{{\'e}}1 {í}{{\'i}}1 {ó}{{\'o}}1 {ú}{{\'u}}1
    {Á}{{\'A}}1 {É}{{\'E}}1 {Í}{{\'I}}1 {Ó}{{\'O}}1 {Ú}{{\'U}}1
    {à}{{\`a}}1 {è}{{\`e}}1 {ì}{{\`i}}1 {ò}{{\`o}}1 {ù}{{\`u}}1
    {À}{{\`A}}1 {È}{{\'E}}1 {Ì}{{\`I}}1 {Ò}{{\`O}}1 {Ù}{{\`U}}1
    {ä}{{\"a}}1 {ë}{{\"e}}1 {ï}{{\"i}}1 {ö}{{\"o}}1 {ü}{{\"u}}1
    {Ä}{{\"A}}1 {Ë}{{\"E}}1 {Ï}{{\"I}}1 {Ö}{{\"O}}1 {Ü}{{\"U}}1
    {â}{{\^a}}1 {ê}{{\^e}}1 {î}{{\^i}}1 {ô}{{\^o}}1 {û}{{\^u}}1
    {Â}{{\^A}}1 {Ê}{{\^E}}1 {Î}{{\^I}}1 {Ô}{{\^O}}1 {Û}{{\^U}}1
    {œ}{{\oe}}1 {Œ}{{\OE}}1 {æ}{{\ae}}1 {Æ}{{\AE}}1 {ß}{{\ss}}1
    {ç}{{\c c}}1 {Ç}{{\c C}}1 {ø}{{\o}}1 {å}{{\r a}}1 {Å}{{\r A}}1
    {€}{{\EUR}}1 {£}{{\pounds}}1
    %
    {^}{{{\color{ipython_purple}\^{}}}}1
    {=}{{{\color{ipython_purple}=}}}1
    %
    {+}{{{\color{ipython_purple}+}}}1
    {-}{{{\color{ipython_purple}-}}}1
    {*}{{{\color{ipython_purple}$^\ast$}}}1
    {/}{{{\color{ipython_purple}/}}}1
    %
    {+=}{{{+=}}}1
    {-=}{{{-=}}}1
    {*=}{{{$^\ast$=}}}1
    {/=}{{{/=}}}1,
    literate=
    *{-}{{{\color{ipython_purple}-}}}1
     {?}{{{\color{ipython_purple}?}}}1,
    %
    identifierstyle=\color{black}\ttfamily,
    commentstyle=\color{ipython_cyan}\ttfamily,
    stringstyle=\color{ipython_red}\ttfamily,
    keepspaces=true,
    showspaces=false,
    showstringspaces=false,
    %
    rulecolor=\color{ipython_frame},
    frameround={t}{t}{t}{t},
    numbers=none,
    numberstyle=\tiny\color{halfgray},
    %
    %
    backgroundcolor=\color{ipython_bg},
    %   extendedchars=true,
    %basicstyle=\scriptsize,
    basicstyle=\ttfamily\footnotesize,
    columns=fullflexible,
    keywordstyle=\color{ipython_green}\ttfamily,
}

@@

%%% color Comments 
\newcommand{\FrL}[1]{{\color{cyan}FL: #1}}

@@

\begin{document}
\journalinfo{The Open Journal of Astrophysics}
%\submitted{submitted xxx, revised yyy, accepted zzz}

@@

\author{
J.~E. Campagne$^{1,\ast}$, 
F. Lanusse$^2$, 
J. Zuntz$^3$,\\
\jaxcosmo\ contributors}
\thanks{$^\ast$jean-eric.campagne@ijclab.in2p3.fr}

\affiliation{$^1$Université Paris-Saclay, CNRS/IN2P3, IJCLab, 91405 Orsay, France}

%\date{\today}

@@

We show that clear improvements are possible using HMC compared to Metropolis-Hasting, and that the Normalizing Flows using the Neural Transport is a promising methodology.
{\color{red}
;in the early phases of Bayesian model development JAZ not sure what this means so I left it out}

@@

%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle

@@

In this context we have developed the open source \texttt{jax-cosmo} library \footnote{\url{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}}, which we present in this paper. The package represents a first step in making the powerful features described above useful for cosmology; it implements a selection of theory predictions for key cosmology observables as differentiable JAX functions.

@@

\FrL{Add description of problem and small demos}

@@

%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
def likelihood(cosmo):
  # Compute mean and covariance of angular Cls, 
  # for specific probes
  mu, cov = gaussian_cl_covariance_and_mean(cosmo, 
                    ell, probes)
  # Return likelihood value
  return gaussian_log_likelihood(data, mu, cov)

@@

# Compute Fisher matrix of cosmological 
# parameters
F = -jax.hessian(likelihood)(cosmo)
\end{lstlisting}
%\end{minted}

@@

Fisher forecasts are also commonly used in survey and analysis strategy strategy, where running a full posterior analysis for each possible choice would be unfeasible. The inverse area of a projection of a Fisher matrix in parameters of interest can be used as a metric for survey constraining power, such as in the Dark Energy Task Force report \citep{2006astro.ph..9591A}.

@@

-\mathcal{L}(\lambda) = \underbrace{\mathbb{E}_{\zeta\sim \mathcal{N}(0,I)}\left[ \log p(T^{-1}(S_\lambda^{-1}(\zeta)),\mathcal{D}) + \log |J_{T^{-1}}(S_\lambda^{-1}(\zeta))| \right]}_{model} \\ + \underbrace{\mathbb{H}[q(\xi;\lambda)]}_{guide}

@@

to get $\nabla_\lambda \mathcal{L}$, the $\nabla$ operator can be put inside the expectation which leads to\footnote{To simplify the notation, $T^{-1}(S_\lambda^{-1}(\zeta))$ has been replaced by $z$.}

@@

Results in terms of sampling efficiency are shown on Tab.~\ref{tab-ESS-NUTS_SVI-1} showing that the SVI followed by Neutral Transport gives high efficiency at low number of samples which is quite interesting to ease the model development in its early phase.
%
\section{Discussion}
\label{sec-discussion}
%
The key question is: what is the HMC/NUTS gain compared to the well known highly optimized \texttt{Cobaya} sampler? 
In one hand, one may argue considering the elapse time of execution that the number of efficient samples per model call is a right metric to consider: 
\begin{equation}
    \eta = \frac{n_{eff}}{\#\ of\ calls} = \frac{N_s \times \varepsilon}{N_s \times n_{steps}} = \frac{\varepsilon}{n_{steps}}
\end{equation}
with $N_s$ the total number of samples, $\varepsilon$ the sampler efficiency and $n_{steps}$ the number of steps (calls) per sample generation. Then, for \texttt{Cobaya} one roughly gets $\varepsilon\approx 3\%$ with $n_{steps}=1$ while for the NUTS sampler $\varepsilon\approx 50\%$ but at the expense of $n_{steps}=2^5$ or more. At the end of the day after tuning generation parameters, the $\eta$ values for both methods would certainly be comparable (ie. about $1\%$). So, at first sight one may conclude that there is no gain to use HMC/NUTS compared to standard MCMC sampler.

@@

\FrL{mention something about scaling with parallel MCMC}

@@

%%%
\section*{CRediT authorship contribution statement}
\textbf{J.E Campagne} Conceptualization, Methodology, Software, Validation, Writing.
\textbf{F. Lanusse}, \textbf{J. Zuntz}...

%\section*{Declaration of Competing Interest}
%The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

@@

Then, using the \numpyro\ PPL, we have set up a \textit{forward model} following the reference \citep{2018PhRvD..98d3526A}. The prior descriptions of the different model parameters are described in Table \ref{tab-DESY1}.  For instance the $\Omega_c$ parameter is treated as random variable as followed
%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
Omega_c = numpyro.sample('Omega_c', Uniform(0.1, 0.9))
\end{lstlisting}
%\end{minted}

Then, we generate mock angular power spectra $C_\ell$ from the auto \& cross correlations of the Number Counts and Weak Lensing probes computed with the \texttt{gaussian\_cl\_covariance\_and\_mean} function in the \texttt{jc.angular\_cl} module. One reads
%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
# Define the lensing and number counts probe
probes = [jc.probes.WeakLensing(nzs_s_sys, 
                       ia_bias=b_ia,
                       multiplicative_bias=m),
          jc.probes.NumberCounts(nzs_l, b)]
cl, C = gaussian_cl_covariance_and_mean(cosmo, 
                 ell, probes, 
                 f_sky=0.25, sparse=True)
\end{lstlisting}
%\end{minted} 
with \texttt{cosmo} an instance of the \texttt{jc.Cosmology} setting the cosmological parameters generated with the priors, and \texttt{ell} (ie. $\ell$) a series of 50 angular modes. The generation of the mock data yields from the possibility to execute a conditioned model with a fiducial cosmology, leaving \numpyro\ dealing with the random number generation:
%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
fiducial_model = condition(model,
    {'Omega_c':0.2545, 'sigma8':0.801, 
    'h':0.682, 'Omega_b':0.0485, 'w0':-1.,'n_s':0.971,
     'A':0.5,'eta':0.,
     'm1':0.0,'m2':0.0,'m3':0.0,'m4':0.0,
     'dz1':0.0,'dz2':0.0,'dz3':0.0,'dz4':0.0,
     'b1':1.2,'b2':1.4,'b3':1.6,'b4':1.8,'b5':2.0
      })

with seed(rng_seed=42):
    data, P, C = fiducial_model()
\end{lstlisting}
%\end{minted}

The \textit{inference model} looks very similar to the \textit{forward model} excepts that first we fix the angular power spectra covariance matrix $C$ (and its inverse $P$) and secondly we let \numpyro\ generates the model $C_{\ell}$ from the priors
%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
cl = jc.angular_cl.angular_cl(cosmo, ell, 
                              probes).flatten()
return numpyro.sample('cl', MultivariateNormal(cl, 
                            precision_matrix=P,
                            covariance_matrix=C))
\end{lstlisting}
%\end{minted}

These theoretical $C_{\ell}$ are in turn conditioned by the mock $C_\ell$ as followed
%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
observed_model = condition(model, {'cl': data})
\end{lstlisting}
%\end{minted}
Then, as an example the optimisation of the cosmological parameters using for instance the NUTS flavour of MCMC can be conducted as followed
%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
nuts_kernel = numpyro.infer.NUTS(observed_model,
              step_size=1e-1, 
              init_strategy=numpyro.infer.init_to_median,
              dense_mass=True,
              max_tree_depth=7)

@@

mcmc.run(jax.random.PRNGKey(42))
\end{lstlisting}
%\end{minted}
More details can be found in the repository of this paper at this link \nblink{VectorizedNumPyro_block}.
%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%\section*{Bibliography}
%\bibliographystyle{elsarticle-harv} 
\bibliographystyle{aa} 
\typeout{}
\bibliography{refs}

>>>>

%% 
%% Copyright 2007-2019 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references

%%\documentclass[preprint,12pt,authoryear]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times,authoryear]{elsarticle}
%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,3p,times,authoryear]{elsarticle}
%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,5p,times,authoryear]{elsarticle}
\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{minted}
\usepackage{amssymb}
\usepackage{amsmath}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}
\usepackage{aas_macros}
\usepackage{fontawesome}
%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\usepackage{xcolor}
\definecolor{xlinkcolor}{cmyk}{1,1,0,0}
\usepackage[colorlinks = true,
            linkcolor = xlinkcolor,
            urlcolor  = xlinkcolor,
            citecolor = xlinkcolor,
            anchorcolor = xlinkcolor]{hyperref}

@@

\newcommand{\FL}[1]{{\color{cyan}FL: #1}}

\journal{Astronomy and Computing}

@@

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

@@

\author{\jaxcosmo\ contributors}

\address{}

@@

We show that clear improvements are possible using HMC compared to Metropolis-Hasting, and that the Normalizing Flows using the Neural Transport is a promising methodology.
;in the early phases of Bayesian model development JAZ not sure what this means so I left it out

@@

%%Graphical abstract
%\begin{graphicalabstract}
%\includegraphics{grabs}
%\end{graphicalabstract}

%%Research highlights
%\begin{highlights}
%\item Research highlight 1
%\item Research highlight 2
%\end{highlights}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

@@

In this context we have developed the open source \texttt{jax-cosmo} library \footnote{ \url{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}}, which we present in this paper. The package represents a first step in making the powerful features described above useful for cosmology; it implements a selection of theory predictions for key cosmology observables as differentiable JAX functions.

@@

\FL{Add description of problem and small demos}

@@

\begin{minted}[fontsize=\footnotesize]{python}
def likelihood(cosmo):
  # Compute mean and covariance of angular Cls, 
  # for specific probes
  mu, cov = gaussian_cl_covariance_and_mean(cosmo, 
                    ell, probes)
  # Return likelihood value
  return gaussian_log_likelihood(data, mu, cov)

@@

# Compute Fisher matrix of cosmological 
# parameters
F = - jax.hessian(likelihood)(cosmo)
\end{minted}

@@

Fisher forecasts are also commonly used in survey and analysis strategy strategy, where running a full posterior analysis for each possible choice would be unfeasible. The inverse area of a projection of a Fisher matrix in parameters of interest can be used as a metric for survey constraining power, such as in the Dark Energy Task Force report \citep{detf}.

@@

-\mathcal{L}(\lambda) = \underbrace{\mathbb{E}_{\eta\sim \mathcal{N}(0,I)}\left[ \log p(T^{-1}(S_\lambda^{-1}(\zeta)),\mathcal{D}) + \log |J_{T^{-1}}(S_\lambda^{-1}(\zeta))| \right]}_{model} \\ + \underbrace{\mathbb{H}[q(\xi;\lambda)]}_{guide}

@@

to get $\nabla_\lambda \mathcal{L}$, the $\nabla$ operator can be put inside the expectation which leads to

@@

Results in terms of sampling efficiency are shown on Tab.~\ref{tab-ESS-NUTS_SVI-1} showing that the SVI followed by Neutral Transport gives high efficiency at low number of samples which is quite interesting to use these schema to ease the model development in its early phase.
%
\section{Discussion}
\label{sec-discussion}
%
The key question is: what is the HMC/NUTS gain compared to the well known highly optimized \texttt{Cobaya} sampler? 
In one hand, one may argue considering the elapse time of execution that the number of efficient samples per model call is a right metric to consider: 
\begin{equation}
    \eta = \frac{n_{eff}}{\#\ of\ calls} = \frac{N_s \times \varepsilon}{N_s \times n_{steps}} = \frac{\varepsilon}{n_{steps}}
\end{equation}
with $N_s$ the total number of samples, $\varepsilon$ the sampler efficiency and $n_{steps}$ the number of steps (calls) per sample generation. Then, for \texttt{Cobaya} one roughly gets $\varepsilon\approx 3\%$ with $n_{steps}=1$ while for the NUTS sampler $\varepsilon\approx 50\%$ but at the expense of $n_{steps}=2^5$ or more. At the end of the day after tuning generation parameters, the $\eta$ values for both methods would certainly be comparable (ie. about $1\%$). So, at first sight one may conclude that there is no gain to use HMC/NUTS compared to standard MCMC sampler.

@@

\FL{mention something about scaling with parallel MCMC}

@@

%%%

{\color{red} the next sections are written along the model of \url{https://www.sciencedirect.com/science/article/pii/S2213133722000373} in Astronomy and Computing}
%%%
\section*{CRediT authorship contribution statement}
\textbf{F. Lanusse}, \textbf{J. Zuntz}...
\textbf{J.E Campagne} Conceptualization, Methodology, Software, Validation, Writing.

\section*{Declaration of Competing Interest}
The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

@@

Then, using the \numpyro\ PPL, we have set up a \textit{forward model} following the reference \citep{2018PhRvD..98d3526A}. The prior descriptions of the different model parameters are described in Table \ref{tab-DESY1}.  For instance the $\Omega_c$ parameter is treated as random variable as followed
\begin{minted}[fontsize=\footnotesize]{python}
Omega_c = numpyro.sample('Omega_c', Uniform(0.1, 0.9))
\end{minted}
Then, we generate mock angular power spectra $C_\ell$ from the auto \& cross correlations of the Number Counts and Weak Lensing probes computed with the \texttt{gaussian\_cl\_covariance\_and\_mean} function in the \texttt{jc.angular\_cl} module. One reads
\begin{minted}[fontsize=\footnotesize]{python}
# Define the lensing and number counts probe
    probes = [jc.probes.WeakLensing(nzs_s_sys, 
                                    ia_bias=b_ia,
                       multiplicative_bias=m),
             jc.probes.NumberCounts(nzs_l, b)]
cl, C = gaussian_cl_covariance_and_mean(cosmo, ell, probes, 
                               f_sky=0.25, sparse=True)
\end{minted} 
with \texttt{cosmo} an instance of the \texttt{jc.Cosmology} setting the cosmological parameters generated with the priors, and \texttt{ell} (ie. $\ell$) a series of 50 angular modes. The generation of the mock data yields from the possibility to execute a conditioned model with a fiducial cosmology, leaving \numpyro\ dealing with the random number generation:
\begin{minted}[fontsize=\footnotesize]{python}
fiducial_model = condition(model,
    {'Omega_c':0.2545, 'sigma8':0.801, 
    'h':0.682, 'Omega_b':0.0485, 'w0':-1.,'n_s':0.971,
     'A':0.5,'eta':0.,
     'm1':0.0,'m2':0.0,'m3':0.0,'m4':0.0,
     'dz1':0.0,'dz2':0.0,'dz3':0.0,'dz4':0.0,
     'b1':1.2,'b2':1.4,'b3':1.6,'b4':1.8,'b5':2.0
      })

with seed(rng_seed=42):
    data, P, C = fiducial_model()
\end{minted}
The \textit{inference model} looks very similar to the \textit{forward model} excepts that first we fix the angular power spectra covariance matrix $C$ (and its inverse $P$) and secondly we let \numpyro\ generates the model $C_{\ell}$ from the priors
\begin{minted}[fontsize=\footnotesize]{python}
cl = jc.angular_cl.angular_cl(cosmo, ell, probes).flatten()
return numpyro.sample('cl', MultivariateNormal(cl, 
                            precision_matrix=P,
                            covariance_matrix=C))
\end{minted}
These theoretical $C_{\ell}$ are in turn conditioned by the mock $C_\ell$ as followed
\begin{minted}[fontsize=\footnotesize]{python}
observed_model = condition(model, {'cl': data})
\end{minted}
Then, as an example the optimisation of the cosmological parameters using for instance the NUTS flavour of MCMC can be conducted as followed
\begin{minted}[fontsize=\footnotesize]{python}
nuts_kernel = numpyro.infer.NUTS(observed_model,
                     step_size=1e-1, 
                     init_strategy=numpyro.infer.init_to_median,
                     dense_mass=True,
                     max_tree_depth=7)

@@

mcmc.run(jax.random.PRNGKey(42))
\end{minted}
More details can be found in the repository of this paper at this link \nblink{VectorizedNumPyro_block}.
%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%\section*{Bibliography}
\bibliographystyle{elsarticle-harv} 
\typeout{}
\bibliography{refs}
